---
title: "Assignment_3_krill"
author: "Erika Scheibe"
format: 
  html:
    self-contained: true
    code-tools: 
      source: true
editor: visual
date: today
execute:
  warning: false
  message: false
---

```{r}

#| label: load-packages
#| message: false

library(tidyverse)
library(here)
library(data.table)
library(janitor)
library(readxl)
library(ggfortify)
library(conflicted)
conflict_prefer("filter", winner = "dplyr")
library(broom)
library(car)

```

```{r}
#open file


#| label: Import and clean data
#| message: false

filebehere <- here("data/Schaafsma-etal_2021.tab")

krilll <- read_delim(
  file = filebehere,
  delim = "\t",
  escape_double = FALSE,
  trim_ws = TRUE,
  skip = 192) |> 
   janitor::clean_names() |> 
  dplyr::select(Species = species,
    Total_length_mm = tl_mm_see_comment_stereo_microscope,
    Wet_mass_m = wet_m_mg_total_wet_mass_of_individual) |> 
dplyr::filter(Species == "Euphausia superba") |> 
dplyr::mutate(
    Log10_Total_length_mm = log10(Total_length_mm),
    Log10_Wet_mass_m = log10(Wet_mass_m))

```

```{r}

#| label: Suitable reg model, Diagnostic plots
#| message: false

mod <- lm(Log10_Wet_mass_m ~ Log10_Total_length_mm, data = krilll)
mod

autoplot(mod) 

#predict(object = mod)
```

I checked for non-normal distribution of the residuals. The points are on a straight line for the most part on the QQ plot; it appears to be normally distributed. There are a few outliers, but not that many. These outliers should be investigated to understand why they may be outliers (just a rare-sized krill, or incorrect data entry maybe (if number is way outside the norm), etc). There are not many outliers so I'll include it in the analysis.

```{r}
#| label: Expected wet mass (in mg) and prediction interval of a 30mm long E. superba
#| message: false

augment(mod, interval = "prediction") |>
  ggplot(aes(x = Log10_Total_length_mm, y = Log10_Wet_mass_m)) +
  geom_point() +
  theme_bw() +
  geom_ribbon(aes(ymin = .lower, ymax = .upper, alpha = .3)) +
  geom_line(aes(y = .fitted)) +
  labs(x = "Log10 Total E. superba length in mm",
       y = "Log10 Total wet mass of individuals in mg") +
  ggtitle("Allometry model: Euphausia superba prediction") +
  geom_vline(xintercept = log10(30), linetype = "dotted")
```

Model results

The least-squares regression model assumptions for body allometry or relationship between log-transformed total length and log-transformed wet mass for Euphausia superba were met. The model predicted that the expected wet mass of a single 30 mm long E. superba was 179.5 mg wet mass with a 95% confidence interval range of 175.6 to 183.4. The model gave significant results (p \< 0.001). 99% of the variance was explained by the predictor.
